Diversification of Australian blind snakes

### 2022-12-22 manuscript.Rmd
upudate manuscript

Consistent with previous research by @marinTracing2013, we found that the initial diversification of Anilios occurred in the Tropical Grassland biome, followed by transitions and/or expansion into other biomes. However, our analyses indicated that species-level divergence occurred earlier, between 20 – 10 MYA, compared to the 15 – 10 MYA proposed by @marinTracing2013. Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008]. This indicates that many species found in arid biomes, particularly in Group 2, diversified in the arid biome prior to the onset of aridification in the late Miocene [@byrneBirth2008].

@@

ALEX SKEELS add to this discussion?
- Paleoclimate reconstructions based on DEM by Scotese and Straume were similar in profile. However, DEM base on Valdes differed from the other two. Scotese is characterised by repeated fluctuations in the Pleistocene to reflect glacial cycles. to add discussion to this?

@@

- Diversification rate decrease with aridification. But this is not explained by lineages transitioning into the arid biome as diversification rates did not differ between lineages in arid and non-arid biomes. Turnover rate also ...

- Lineages that have evolved more recently in the arid zone became extinct with the Plio-Pleistocene glacial cycles leaving only species that have survived in pockets of refugia. 
- Vast arid biome covers 70% of the continent, which the largest biome in Australia with lots of heterogeneity [@pepperLife2021]. 
- Pockets of certain areas were refugia during glaciation cycles of the Pleistocene.

- Warmer and dryer environment may not directly decrease speciation rate because blind snakes are physically adapting to warmer temperatures. However, climatic fluctuations can contract and fragment available niche for blind snakes and thus influence diversification rates.

@@

- Alternatively, since they are fossorial they may not be subject to the same selective pressures epigean (surface dwelling) species. For example, their association with their prey, ants and termites, may buffer them from dessication and predation pressures (Nevo 1979** cite)

- Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].

@@

@deepakAridification2018 - accelerated speciation with aridification from the late Miocene. Probably due to emergence of open habitats
@catulloAridification2014 - multiple independent origing from mesic or monsoonal ancestors into the arid zone. Lineages older than intense aridification. 
@fujitaDiversification2010 - independent origins of arid zone lineages in _Heteronotia_
@pepperGenetic2011 - mountain ranges harbour more diversity than deserts. arid zone vertebrates show little genetic diff. But recent population expansion in the desert. 
@oliverIndependent2014 - multiple independent invasion into the arid zone.
@owenHow2017 - cicada increase speciation in arid zone compared to mesic zone (SSE)
@toussaintUnveiling2015 - diving beetle boom and bust diversification. Invaded arid zone and changed ecologically

>>>>

Consistent with previous research by @marinTracing2013, we found that the initial diversification of _Anilios_ occurred in the Tropical Grassland biome, followed by transitions and/or expansion into other biomes. However, our analyses indicated that species-level divergence occurred earlier, between 20 – 10 MYA, compared to the 15 – 10 MYA proposed by @marinTracing2013. Our BioGeoBEARS analyses suggest that _Anilios_ gradually diversified in the tropical grassland (Group 1) and arid biome (Group 2) during a time when Australia was relatively wet and humid [@byrneBirth2008]. This indicates that many species found in arid biomes, particularly in Group 2, diversified in the arid biome prior to the onset of aridification in the late Miocene [@byrneBirth2008].

@@

ALEX SKEELS add to this discussion?
- Paleoclimate reconstructions based on DEM by Scotese and Straume were similar in profile. However, DEM base on Valdes differed from the other two. Scotese is characterised by repeated fluctuations in the Pleistocene to reflect glacial cycles.

@@

Several studies have shown that arid-adapted species originated either from a single ancestor present in Australia before the Miocene, or through multiple independent expansion into the arid zone [@byrneBirth2008]. Our ancestral biome analyses show that several non-arid ancestors of _Anilios_ blind snakes have diverged into the arid zone evident by multiple arid species throughout the phylogeny. This pattern is consistent with several other groups vertebrates that started to diversify in the Miocene including goannas [@brennanPhylogenomicsMonitorLizards2021;@pavon-vazquezCompetition2022], frogs [@catulloAridification2014], and several geckos [@fujitaDiversification2010;@oliverIndependent2014].

Although the expansion of the arid zone and fragmentation of mesic environments throughout the Miocene has been linked with increase speciation [cite], studies have found that mountainous areas in the arid zone habour more diversity than the flat featureless desert [@pepperPalaeoclimate2011].

-- new open dry habitats for different species of reptiles to fill [@sandersMolecular2008;@hugallPhylogeny2008], which

our state speciation-extinction analyses on the blind snake phylogeny also indicate that speciation, extinction, and turnover rates did not differ between lineages in arid and non-arid biomes.

- Warmer and dryer environment may not directly decrease speciation rate because blind snakes are physically adapting to warmer temperatures. However, climatic fluctuations can contract and fragment available niche for blind snakes and thus influence diversification rates. 
- Alternatively, since they are fossorial they may not be subject to the same selective pressures epigean (surface dwelling) species. For example, their association with their prey, ants and termites, may buffer them from dessication and predation pressures (Nevo 1979** cite)
- Microhabitat better explains diversification than climate, but typhlopidae has the second highest speciation rates among fossorial snakes [@bars-closelDiversification2017].

- Lineages that have evolved more recently in the arid zone became extinct with the Plio-Pleistocene glacial cycles leaving only species that have survived in pockets of refugia. 
- Pockets of certain areas were refugia during glaciation cycles of the Pleistocene.

@@

@@

<!---
@pepperGenetic2011 - mountain ranges harbour more diversity than deserts. arid zone vertebrates show little genetic diff. But recent population expansion in the desert. 
@catulloAridification2014 - multiple independent origin from mesic or monsoonal ancestors into the arid zone. Lineages older than intense aridification. 
@fujitaDiversification2010 - independent origins of arid zone lineages in _Heteronotia_
@oliverIndependent2014 - multiple independent invasion into the arid zone.
@owenHow2017 - cicada increase speciation in arid zone compared to mesic zone (SSE)
@toussaintUnveiling2015 - diving beetle boom and bust diversification. Invaded arid zone and changed ecologically
@deepakAridification2018 - India accelerated speciation with aridification from the late Miocene. Probably due to emergence of open habitats
--->
